Nitric oxide feedback to ciliary photoreceptor cells gates a UV avoidance circuit

### 2023-06-08 Manuscript/Manuscript.Rmd
Upadate Supplement files
Ms: methods edits
Ms: methods edit

We assayed larval photoresponses in a vertical Plexiglas column (31 mm x 10 mm x 160 mm water height).
For stimulation, we illuminated the column from the top with a monochromator (Polychrome II, Till Photonics) controlled by AxioVision 4.8.2.0 (Carl Zeiss MicroImaging GmbH) via analog voltage.
The light passed a collimator lens (LAG-65.0-53.0-C with MgF2 Coating, CVI Melles Griot) before entering the column.
The column was illuminated from both sides with light-emitting diodes (LEDs).
The LEDs on each side were grouped into two strips.
One strip contained UV (395 nm) LEDs (SMB1W-395, Roithner Lasertechnik) and the other infrared (810 nm) LEDs (SMB1W-810NR-I, Roithner Lasertechnik).
The UV LEDs were run at 4 V. The infrared LEDs were run at 8 V (overvoltage) to illuminate the larvae for imaging with a DMK camera (DMK 22BUC03, The Imaging Source).
We recorded videos at 15 frames per second with the software IC Capture (The Imaging Source).

@@

The synthetic peptides contained an N-terminal Cys that was used for coupling during purification.
Antibodies were affinity purified from sera as previously described [@Conzelmann2012].

@@

The *NOS* promoter (fragment sizes: 12 Kb) was amplified and cloned upstream of 3xHA- Palmi-tdTomato.
Larvae injected with promoter constructs (ca. 250 ng/ml) were analysed for reporter expression at 3 days post fertilization using an AxioImager Z.1 fluorescence wide-field microscope (Carl Zeiss GmbH, Jena) and immediately fixed for immunostainings.
The protocol followed for immunostaining of HA-tagged reporters was recently described [@veraszto2017].
Specimens were imaged with a LSM 780 NLO or LSM 880 with Airysan Confocal Microscope (Zeiss, Jena).

@@

For calcium imaging, 49-55 hpf larvae were used.
Experiments were carried out at room temperature and larvae were immobilised by being embedded in 2.5% agarose filtered artificial seawater between a slide and coverslip spaced with adhesive tape.
GCaMP6s mRNA (1 mg/ml) was injected into zygotes as described previously [@Randel2014a].
Larvae were imaged on a Zeiss LSM 880 with Airyscan (with a C-Apochromat 63X/1.2 Corr - water) with a frame rate of 1.88 frame/sec and an image size of 512 x 512 pixels.
The larvae were stimulated in a region of interest (a circle with 50 pixel diameter) with 405 nm lasers controlled by the Bleaching mode.
The imaging laser had a similar intensity than the stimulus laser but covered an area that was 10 times larger than the stimulus ROI.

**Cell culture experiment**

Green cGull was used for the cGMP assay [@Matsuda2017].
A full-length Pdum-NIT-GC1 and -NIT-GC2 coding sequences were amplified by PCR starting from a *Platynereis dumerilii* cDNA library and cloned into the pcDNA3.1(+) vector using the T2A self-cleaving sequence.
COS-7 cells with low expression of endogenous soluble guanylate cyclase were used as cultured cells for gene expression.
This cell line was purchased from Angio-proteomie (CAT no. cAP-0203).
The COS-7 cells were maintained at 37 °C in 35mm dishes (Nunc™ Glass Bottom Dishes) containing 3 mL of DMEM, high glucose glutamax medium (Thermo; Cat. No. 10566016) supplemented with 10% fetal bovine serum (Thermo; Cat. No. 10082147).
Upon reaching confluency of approximately 85%, we transfected the cells with the plasmid containing Green cGull-T2A-NITGC1.
Transfections were carried out with 150 ng of each plasmid and 0.3 μl of the transfection Lipofectamine 3000 Reagent (invitrogen; Cat. No. L3000001).
Two days post-transfection, we removed the culture medium and substituted it for fresh DMEM-medium.
For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5mM NaH2PO4 , 0.5mM S-3 MgSO4 , 1.5 mM CaCl2 , 10 mM HEPES, 2 mM NaHCO3 and 5 mM glucose).
Dishes were mounted on a stage heated at 37 °C and imaging was performed using an inverted microscope (LSM880, Zeiss) equipped with an oil-immersion objective lens (UApo/340, 40, NA = 0.17).
Images were acquired using a xenon lamp, 460--495 nm excitation filter, 505-nm dichroic mirror and 510-- 550-nm emission filter (Zeiss).
S-Nitroso-N-acetyl-D, L-penicillamine (SNAP) was purchased from Sigma-Aldrich (St. Louis, MO, USA) .
The exposure time of the EM-CCD camera was controlled by the ZEN software (Zeiss).
Images were acquired every 15 s for 10 min and stimulation was initiated 2 min after starting image acquisition.
Imaging data analysis was performed using ImageJ (National Institutes of Health, Bethesda, MD, USA).

>>>>

We assayed larval photoresponses in a vertical Plexiglas column (31 mm x 10 mm x 160 mm water height).
For stimulation, we illuminated the column from the top with a monochromator (Polychrome II, Till Photonics) controlled by AxioVision 4.8.2.0 (Carl Zeiss GmbH, Jena) via analog voltage.
The light passed a collimator lens (LAG-65.0-53.0-C with MgF2 Coating, CVI Melles Griot) before entering the column.
The column was illuminated from both sides with light-emitting diodes (LEDs).
The LEDs on each side were grouped into two strips.
One strip contained UV (395 nm) LEDs (SMB1W-395, Roithner Lasertechnik) and the other infrared (810 nm) LEDs (SMB1W-810NR-I, Roithner Lasertechnik).
The UV LEDs were run at 4 V. The infrared LEDs were run at 8 V (overvoltage) to illuminate the larvae for imaging with a DMK camera (DMK 22BUC03, The Imaging Source).
We recorded videos at 15 frames per second with the software IC Capture (The Imaging Source).

@@

Antibodies were raised in rats or rabbits against synthetic peptides containing an N-term Cys residue (Altabioscience). The same Cys-containing synthetic peptides were used for affinity purification.
Antibodies were affinity purified from sera on a SulfoLink Coupling Resin (Thermo Fischer) as previously described [@Conzelmann2012].

@@

A 12 kb genomic region upstream of the start site of the *NOS* gene was amplified and cloned upstream of 3xHA-Palmi-tdTomato to yield the plasmid *NOS-3xHA-Palmi-tdTomato*.
Larvae injected with the promoter construct (ca. 250 ng/ml) were analysed for reporter expression at three days post fertilization on an AxioImager Z.1 fluorescence microscope (Carl Zeiss GmbH, Jena) and if positive fixed for immunostaining.
The immunostaining for the HA-tagged reporter was done as described [@veraszto2017].
Stained specimens were imaged on an LSM 780 NLO or LSM 880 Airysan Confocal Microscope (Carl Zeiss GmbH, Jena).

@@

For calcium imaging, 49-55 hours post fertilisation larvae were used.
Experiments were carried out at room temperature and larvae were immobilised in 2.5% agarose in artificial sea water and mounted between a slide and coverslip spaced with adhesive tape.
GCaMP6s mRNA (1 mg/ml) was injected into zygotes as described previously [@Randel2014a].
Larvae were imaged on a Zeiss LSM 880 with Airyscan (with a C-Apochromat 63X/1.2 Corr - water) with a frame rate of 1.88 frame/sec and an image size of 512 x 512 pixels.
The larvae were stimulated in a region of interest (a circle with 50 pixel diameter) with a 405 nm laser controlled by the Bleaching mode of the Zeiss software.
The imaging laser had a similar intensity than the stimulus laser but covered an area that was 10 times larger than the stimulus ROI.

**cGMP production assay**

To measure cGMP production in a cell culture assay we used the Green cGull reporter [@Matsuda2017].
Full-length *NIT-GC1* and *NIT-GC2* coding sequences were amplified by PCR from a *Platynereis dumerilii* cDNA library and cloned into the pcDNA3.1(+) vector with the T2A self-cleaving sequence.
For the assay, COS-7 cells (Angio-proteomie, CAT no. cAP-0203) with a low endogenous level of soluble guanylate cyclase activity were used.
Cells were maintained at 37 °C in 35 mm dishes (Nunc™ Glass Bottom Dishes) containing 3 ml of DMEM, high glucose glutamax medium (Thermo; Cat. No. 10566016) supplemented with 10% fetal bovine serum (Thermo; Cat. No. 10082147).
Upon reaching approximately 85% confluency, cells were transfected with the pcDNA3.1(+)-Green-cGull-T2A-NITGC1 plasmid.
Transfections were carried out with 150 ng of each plasmid and 0.3 μl of Lipofectamine 3000 Reagent (invitrogen; Cat. No. L3000001).
Two days post-transfection, the culture medium was substituted for fresh DMEM-medium. 
For single-wavelength imaging experiments, cells in 35-mm dishes were washed twice and imaged in modified Ringer's buffer (140 mM NaCl, 3.5 mM KCl, 0.5 mM NaH~2~PO~4~ , 0.5 mM S-3 MgSO~4~, 1.5 mM CaCl~2~ , 10 mM HEPES, 2 mM NaHCO~3~ and 5 mM glucose).
Dishes were mounted on a stage heated to 37 °C and imaging was done on an inverted confocal microscope (LSM880, Carl Zeiss GmbH, Jena) equipped with an oil-immersion objective lens (UApo/340, 40x, NA = 0.17).
Images were acquired using a xenon lamp, a 460-495 nm excitation filter, a 505-nm dichroic mirror and a 510-550-nm emission filter (Carl Zeiss GmbH, Jena).
The exposure time of the EM-CCD camera was controlled by the ZEN software (Carl Zeiss GmbH, Jena).
Images were acquired every 15 s for 10 min and stimulation was initiated 2 min after starting image acquisition.
Imaging data analysis was done with ImageJ (National Institutes of Health, Bethesda, MD, USA).
As NO donor, we used S-Nitroso-N-acetyl-D, L-penicillamine (SNAP, Sigma-Aldrich, St. Louis, MO, USA).

## Data analysis and figure preparation

Data were analysed in ImageJ and R.
All figures were assembled in R with the cowplot and patchwork packages.
All scripts are available at <https://github.com/JekelyLab/Joukra_et_al_NOS> (commit ).
